---
title: "Project 3: STAT302package Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Project 3: STAT302package Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Introduction of this package: 
This package is created as a class project for Stat302 in UW. It contains four formulas: my_t_test, my_lm, my_knn_cv, 
and my_rf_cv. This package is collabrated by Celeste Zeng and Simona Liao

## How to install the package from github:
```{r, eval=FALSE}
devtools::install_github("celestezeng33/Stat302package", build_vignette = TRUE, build_opts = c())
library(Stat302package)
# Use this to view the vignette in the Demo HTML help
help(package = "Stat302package", help_type = "html")
# Use this to view the vignette as an isolated HTML file
utils::browseVignettes(package = "Stat302package")
```

```{r setup}
library(STAT302package)
library(ggplot2)
library(magrittr)
```
A tutorial for the function my_knn_cv in the STAT302packgePredict 
Using the my_gapminder dataset, we predict output class continent using covariates gdpPercap and lifeExp by utilizing 5-fold cross validation (k_cv = 5).

```{r}
tu_knn <- my_knn_cv(train = my_gapminder[, 3 : 4], 
          cl = my_gapminder$continent, k_nn = 10, k_cv = 5)
```

```{r}
k_vector <- c(2, 5, 10)
tu_rf <- matrix(NA, nrow = 90, ncol = 2)
for (i in 1 : length(k_vector)) {
  for (j in 1 : 30) {
    tu_rf[(i - 1) * 30 + j, 1] <- k_vector[i]
    tu_rf[(i - 1) * 30 + j, 2] <- my_rf_cv(k = k_vector[i])
  }
}
```

```{r}
tu <- data.frame("K_value" = tu_rf[, 1], "MSE" = tu_rf[, 2])
tu_graph <- ggplot(data = tu,
       aes(x = K_value, y = MSE, group = K_value)) +
  geom_boxplot(fill = "lightblue") +
  theme_bw(base_size = 16) +
  labs(title = "CV estimated MSE for different k-values",
       x = "k-values",
       y = "Estimated MSE") + 
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  scale_x_continuous(breaks = c(2, 5, 10)) 
rf_vector <- c(mean(tu_rf[1 : 30, 2]), mean(tu_rf[31 : 60, 2]), mean(tu_rf[61 : 90, 2]), var(tu_rf[1 : 30, 2]),
               var(tu_rf[31 : 60, 2]), var(tu_rf[61 : 90, 2]))
rf_matrix <- matrix(rf_vector, nrow = 3, ncol = 2)
rf_table <- data_frame("k_values" = c(2, 5, 10),"mean" = rf_matrix[, 1], "variance" = rf_matrix[, 2])
```
